Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-09-29 manuscript.RMD
completed proofreading

CONTINUE PROOFREADING HERE!!!!!

@@

After 5 RT sessions (Figure 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength by 7% more than glucose (Figure 1D, *p* = `r pdiff.c4`). At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (Figure 1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`). Both RT with glucose and placebo led to significantly reduced strength post-fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000). Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`). Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).

@@

From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2). However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.

@@

Overall, RT led to robust increases from pre- to post-intervention in all measured proteins (Figure 3A). RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A). There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).

@@

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA, and rDNA transcription initiation-associated proteins, compared to RT with placebo. Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume. At one time point, RT with glucose led to less reduction in strength compared to placebo. Levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose. Importantly, there were no differences in daily macronutrient intake on pairwise consecutive days. These findings suggest that the effects of high vs. low glucose/glucose starvation condition on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*. Notably, the present study was specifically designed to investigate differences in pre to post changes within participants, and is as such also quite limited to this perspective. Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few. *In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011], and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015]. These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015]. In addition, Hillier et al. found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000]. Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo will elicit this effect. Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs. placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Due to the exploratory nature of our hypotheses, this was hardly a very surprising result. Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodent [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare high glucose condition to placebo, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important in supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins.
<!-- OLD: While content of UBF and rpS6 changed equally in RT with glucose compared to placebo, content of c-Myc was significanly higher after RT with placebo. This was unexpected, since the general transcription factor c-Myc previously has been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Along that notion, it would be reasonable to expect that differences in c-Myc content was reflected in total RNA and rRNA. However, this was not the case in the present study. c-Myc's independency from mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally. Moreover, the observation that total RNA and rRNA were unaffected by the significant differences in c-Myc content, does strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. -->The linear relationship found exclusively between UBF content and total RNA accumulation, and neither c-Myc nor rpS6, support the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022]. As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].

As in the biological data, there was generally no difference in the change of strength as an index between the conditions. Generally, skeletal muscle strength decreased from baseline to post-fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at post-fifth session measurement. This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity were quite high. Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11hrs post-exercise, followed by a levelling off or drop until 22hrs post-exercise. Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22hrs [@raastad_recovery_2000]. Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post-sixth RT session and a drop at 23hrs. Further, strength testing during the intervention was conducted 23hrs after RT, meaning that these tests also may have been influenced by the biphasic recovery. However, this does not explain the difference between conditions at post-fifth RT session, or the great drop in strength from post 4RT to post 5RT sessions. A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo. Notably, training volume showed that the total training session volume was equal on pairwise consecutive days, i.e. no difference between days 1-2, days 3-4 and so on. Hence, there were no differences in mechanical loading to induce a higher stress between conditions. Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT <!-- this is possibly a stretch -->[@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999]. Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable. Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design. Initially, we met our minimum limit of 16 participants. Unfortunately, three dropped out during the intervention, leaving our statistical power slightly underpowered according to the a priori calculation. Furhter, we did not measure the intramuscular glycogen storage, representing another limitation to our design. Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo condition. The participants did indeed match their macro nutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose. Thus, it may stand to reason that the participants had similar levels of intramuscular levels before ingesting any supplement and that the glucose supplement led to higher glucose availability during exercise, as by design. However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our outcome measures.

@@

In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults. Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23hrs of the last session. The observations in total RNA, rRNA and rDNA transcription initiation-associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training. If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher-resolution results.

### DATA AVAILABILITY

@@

### SUPPLEMENt MATERIAL

@@

## References

@@

**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization. Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5). Before baseline testing, the participants were randomly allocated to exercise one leg with glucose (GLU) and the other with placebo (PLA), in a unilateral, alternating fashion. Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, the figure illustrates an example where the participant was randomized to start RT with GLU. Biopsies were taken from m. vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2). Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples. Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six session. <!-- A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family - necessary to include this here? Its already reported under "experimental design" -->. B and C) Changes in plasma glucose (B, mmol/L) and serum c-peptide levels (C, pmol/L). Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120). C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT. D) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after two and four RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2). The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations. Values are presented as changes in estimated marginal means ± 95% confidence intervals (CI). * = *p* < 0.05 between groups. Glucose n = 13, placebo n = 13.

**Figure 2:** Changes in total RNA and ribosomal RNA. A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA. Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2. Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two muscle tissue pieces per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA. Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight. Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure illustration. Values are estimated marginal means fold change per leg per supplement ± 95% CI. Glucose n = 13, placebo n = 13.

**Figure 3:** Changes in protein content. A) Fold-change pre-post in levels of protein in placebo and glucose + fold-change difference 95% CI, with representative western blots of the respective proteins. Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g. from gel 3 to 6, as shown by the duplicates (1 and 2). Changes in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel). Mean log fold-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration. Values are estimated marginal means fold change per leg per supplement ± 95% CI. Glucose n = 13, placebo n = 13. B) Representative total protein stain blot. C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate. Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel and total protein per lane factor. Values are presented as log-transformed means.

>>>>

@@

After 5 RT sessions (Figure 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength by 7% more than glucose (Figure 1D, *p* = `r pdiff.c4`). At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (Figure 1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`). Both RT with glucose and placebo led to significantly reduced strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000). Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed a significant mean increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`). Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).

@@

From baseline to after the intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2). However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.

@@

Overall, RT led to robust increases from before to after the intervention in all measured proteins (Figure 3A). RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A). There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).

@@

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA, and rDNA transcription initiation-associated proteins, compared to RT with placebo. Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume. At one time point, RT with glucose led to less reduction in strength compared to placebo. Levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose. Importantly, there were no differences in daily macronutrient intake on pairwise consecutive days. These findings suggest that the effects of high vs. low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*. Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective. Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few. *In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015]. These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015]. In addition, Hillier et al. found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000]. Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect. Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs. placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Due to the exploratory nature of our hypotheses, this was hardly a very surprising result. Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important in supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins. The linear relationship found exclusively between UBF content and total RNA accumulation, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022]. As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].

As in the biological data, there was generally no difference in the change of strength as an index between the conditions. Generally, skeletal muscle strength decreased from baseline to after the fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at the measurement after five sessions. This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity were quite high. Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise. Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22 hours [@raastad_recovery_2000]. Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30 minutes and 2 hours after the sixth RT session and a drop at 23 hours after the sixth session. Further, strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests. However, this does not explain the difference between conditions at observed after five RT session, or the drop in strength from after four RT sessions to after five RT sessions. A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo. Notably, training volume showed that the total training session volume was equal on pairwise consecutive days, i.e. no difference between days 1-2, days 3-4 and so on. Hence, there were no differences in mechanical loading to induce a higher stress between conditions. Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT [@westerblad_mechanisms_1998; @kent-braun_central_1999]. Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999]. Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable. Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design. Initially, we met our minimum limit of 16 participants. Unfortunately, three dropped out during the intervention, leaving our statistical analyses slightly underpowered according to the a priori calculation. Moreover, we did not measure the intramuscular glycogen storage, representing another limitation to our design. Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo conditions. The participants did indeed match their macronutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose. Thus, it may stand to reason that the participants had similar levels of intramuscular levels before ingesting any supplement and that the glucose supplement led to higher glucose availability during exercise, as by design. However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our main outcome measures.

@@

In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults. Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23 hours of the last session. The observations in total RNA, rRNA and rDNA transcription initiation-associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training. If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher-resolution results.

### Data availability

@@

### Supplementary material

@@

\pagebreak
## References

@@

**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization. Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5). Before baseline testing, the participants were randomly allocated to exercise one leg with glucose (GLU) and the other with placebo (PLA), in a unilateral, alternating fashion. Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, the figure illustrates an example where the participant was randomized to start RT with GLU. Biopsies were taken from m. vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2). Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples. Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six session. B and C) Changes in plasma glucose (B, mmol/L) and serum c-peptide levels (C, pmol/L). Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120). C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT. D) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after two and four RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2). The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations. Values are presented as changes in estimated marginal means ± 95% confidence intervals (CI). * = *p* < 0.05 between groups. Glucose n = 13, placebo n = 13.

**Figure 2:** Changes in total RNA and ribosomal RNA. A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA. Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2. Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two muscle tissue pieces per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA. Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight. Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure visualisation. Values are estimated marginal means fold change per leg per supplement ± 95% CI. Glucose n = 13, placebo n = 13.

**Figure 3:** Changes in protein content. A) Fold-change pre-post in levels of protein in placebo and glucose + fold-change difference 95% CI, with representative western blots of the respective proteins. Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g. from gel 3 to 6, as shown by the duplicates (1 and 2). Changes in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel). Mean log fold-change scores of the duplicates were calculated before modelling and reverse-transformed for figure visualisation. Values are estimated marginal means fold change per leg per supplement ± 95% CI. Glucose n = 13, placebo n = 13. B) Representative total protein stain blot. C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate. Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel and total protein per lane factor. Values are presented as log-transformed means.
